### 2020-09-11 paper/perfectGraphs.tex
Pyramids look nice

\paragraph{Finding Pyramids}
\begin{alg}[Test if $G$ contains a Pyramid]
	\label{alg:testPyramid}
	Input: A graph $G$.

	\noindent Output: Returns whether $G$ contains a pyramid as an induced subgraph.
\end{alg}
\begin{algorithmic}[1]
	\mProcedure{Contains-Pyramid}{$G$}
	\mForEach{triangle $b_1, b_2, b_3$}
		\mForEach{$s_1, s_2, s_3$, such that for $1 \leq i < j \leq 3$, $\{b_i, s_i\}$ is disjoint
			\lsx  from $\{b_j, s_j\}$ and $b_ib_j$ is the only edge between them}
			\mIf{there is a vertex $a$, adjacent to all of $s_1, s_2, s_3$, and to at most
				\lsx one of $b_1, b_2, b_3$, such that if $a$ is adjacent to $b_i$, then $b_i = s_i$ \lsx}
				\ls $M \gets V(G) \setminus \{b_1, b_2, b_3, s_1, s_2, s_3\}$
				\mForEach{$m \in M$}
					\ls $S_1(m) \gets$ the shortest path between $s_1$ and $m$ such that
					\lsx $s_2, s_3, b_2, b_3$ have no neighbors in its interior, if such a 
					\lsx path exists.
					\ls calculate $S_2(m), S_3(m)$ similarly
					\ls $T_1(m) \gets$ the shortest path between $m$ and $b_1$, such that
					\lsx $s_2, s_3, b_2, b_3$ have no neighbors in its interior, if such a
					\lsx path exists
					\ls calculate $T_2(m), T_3(m)$ similarly
				\mEndFor
				\mForEach{$m \in M \cup \{b_1\}$}
					\mIf{$s_1 = b_1$}
						\ls $P_1(b_1) \gets$ the one-vertex path with vertex $b_1$
						\ls \algorithmicforeach ~$m \in M$ \algorithmicdo~ $P_1(m) \gets$ \textsc{undefined}
					\mElse
						\ls $P_1(b_1) \gets $ \textsc{undefined}
						\mForEach{$m \in M$}
							\mIf{ $m$ is nonadjacent to all of $b_2, b_3, s_2, s_3$ \AND
								\lsx $S_1(m)$ and $T_1(m)$ both exist \AND
								\lsx $V(S_1(m) \cap T_1(m)) = \{m\}$ \AND
								\lsx there are no edges between $V(S_1(m) \setminus m)$ 
								\lsx and $V(T_1(m) \setminus m)$ 
								\lsx
							}
								\ls $P_1(m) \gets s_1-S_1(m)-m-T_1(m)-b_1$
							\mElse
								\ls $P_1(m) \gets$ \textsc{undefined}
							\mEndIf
						\mEndFor
					\mEndIf
				\mEndFor
				\ls assign $P_2$ and $P_3$ in a similar manner
				\ls $good\_pairs_{1,2} \gets \emptyset$
				\mForEach{$m_1 \in M \cup \{b_1\}$} \Comment{calculate good $(1, 2)$-pairs}
					\mIf{$P_1(m_1) \neq \UND$}
						\ls color black the vertices of $M$ that either belong to
						\lsx $P_1(m_1)$ or have a neighbor in $P_1(m_1)$
						\lsx color all other vertices white.
						\mForEach{$m_2 \in M \cup \{b_2\}$}
							\mIf{$P_2(m_2)$ exists and contains no black vertices}
								\ls add $(m_1, m_2)$ to $good\_pairs_{1,2}$
							\mEndIf
						\mEndFor
					\mEndIf
				\mEndFor
				\ls calculate $good\_pairs_{1,3}$ and $good\_pairs_{2,3}$ in similar way
				\mForEach{triple $m_1, m_2, m_3$ such that $m_i \in M \cup \{b_i\}$}
					\mIf{$\forall_{1\leq i < j \leq 3}$: $(m_i, m_j)$ is a good $(i, j)$-pair}
						\ls \RETURN \TRUE
					\mEndIf
				\mEndFor
			\mEndIf
		\mEndFor
	\mEndFor
	\ls \RETURN \FALSE
	\mEndProcedure
\end{algorithmic}

\begin{algtext2}

	Enumerate all 6-tuples $(b_1, b_2, b_3, s_1, s_2, s_3)$ such that:
	\begin{itemize}
		\item $\{b_1, b_2, b_3\}$ is a triangle
		\item for $1 \leq i < j \leq 3$, $\{b_i, s_i\}$ is disjoint from $\{b_j, s_j\}$ and $b_ib_j$ is the only edge between them
		\item there is a vertex $a$ adjacent to all of $s_1, s_2, s_3$ and to at most one of $b_1, b_2, b_3$, such that if $a$ is adjacent to $b_i$, then $s_i = b_i$.
	\end{itemize}

	There are $O(|V(G)|^6)$ 6-tuples, and for each we check above conditions in $O(|V(G)|)$ time. For each such 6-tuple we follow with the rest of the algorithm.

	Let $M \leftarrow V(G) \setminus \{b_1, b_2, b_3, s_1, s_2, s_3\}$.

	For each $m \in M$, set $S_1(m)$ equal to the shortest path between $s_1$ and $m$ such that $s_2, s_3, b_2, b_3$ have no neighbors in its interior, if such a path exists. Set $S_2$ and $S_3$ similarly. Set $T_1(m)$ to be the shortest path between $m$ and $b_1$, such that $s_2, s_3, b_2, b_3$ have no neighbors in its interior, if such a path exists, and set $T_2$, $T_3$ similarly. It takes $O(|V(G)|^2)$ time to calculate paths $S_i(m)$ and $T_i(m)$ for all $i$ and $m$.

	Now, we will calculate all possible paths $P_i$. For each $m \in M \cup \{b_1\}$ we will define a path $P_1(m)$ with paths $P_2(m)$, $P_3(m)$ defined in a similar manner.

	If $s_1 = b_1$ let $P_1(b_1)$ be the one-vertex path with vertex $b_1$, and let $P_1(m)$ be undefined for each $m \in M$.

	If $s_1 \neq b_1$, then $P_1(b_1)$ is undefined and for all $m \in M$ we will check if all the following are true:
	\begin{itemize}
		\item $m$ is nonadjacent to all of $b_2, b_3, s_2, s_3$
		\item $S_1(m)$ and $T_1(m)$ both exist
		\item $V(S_1(m) \cap T_1(m)) = \{m\}$
		\item there are no edges between $V(S_1(m) \setminus m)$ and $V(T_1(m) \setminus m)$
	\end{itemize}
	If it is true for some $m \in M$, then we assign $P_1(m) \leftarrow s_1-S_1(m)-m-T_1(m)-b_1$, otherwise we let $P_1(m)$ be undefined. It takes $O(|V(G)|^2)$ to check this, for a given $m$. We assign $P_2$ and $P_3$ in a similar manner. Total time of finding all $P_i(m)$ paths for a given 6-tuple is $O(|V(G)|^3)$.

	Now we want to check if there is a triple $m_1, m_2, m_3$, so that $P_1(m_1)$, $P_2(m_2)$, $P_3(m_3)$ form a pyramid. A most obvious approach of enumerating them all would be too slow, so we do it carefully.

	For $1 \leq i < j \leq 3$, we say that $(m_i, m_j)$ is a \emph{good $(i, j)$-pair}, iff $m_i \in M \cup \{b_i\}$, $m_j \in M \cup \{b_j\}$, $P_i(m_i)$ and $P_j(m_j)$ both exist, and the sets $V(P_i(m_i))$,$V(P_j(m_j))$ are both disjoint and $b_ib_j$ is the only edge between them.

	We show how to find the list of all good $(1, 2)$-pairs, with similar algorithm for all other good $(i, j)$-pairs. For each $m_1 \in M \cup \{b_1\}$, we find the set of all $m_2$ such that $(m_1, m_2)$ is a good $(1,2)$-pair as follows.

	If $P_1(m_1)$ does not exist, there are no such good pairs. If it exists, color black the vertices of $M$ that either belong to $P_1(m_1)$ or have a neighbor in $P_1(m_1)$. Color all other vertices white. (We can do this in $O(|V(G)|^2)$) Then for each $m_2 \in M \cup \{b_2\}$, test whether $P_2(m_2)$ exists and contains no black vertices. We do this for all $m_1$ and get a set of all $(1,2)$-good pairs. In similar way we calculate all good $(1,3)$-pair and $(2,3)$-pairs (in $O(|V(G)|^3)$ time).

	Now, we examine all triples $m_1, m_2, m_3$ such that $m_i \in M \cup \{b_i\}$ and test whether $(m_i, m_j)$ is a good $(i, j)$-pair. If we find a triple such that all three pairs are good, we output that G contains a pyramid and stop.

	If after examining all choices of $b_1, b_2, b_3, s_1, s_2, s_3$ we find no pyramid, output that $G$ contains no pyramid. Since there are $O(|V(G)|^6)$ such choices and it takes a time of $O(|V(G)|^3)$ to analyze each one, the total time is $O(|V(G)|^9)$.
\end{algtext2}

It is easy to see, that if algorithm outputs that $G$ contains a pyramid, it indeed does. The proof of the converse is rather technical and we refer to \cite{MC05} for it.

@@

\begin{lemma}
	If $G$ contains configuration of type $\T_2$ then $G$ is not Berge.
\end{lemma}
\begin{proof}
	Let $(v_1, v_2, v_3, v_4, P, X)$ be a configuration of type $\T_2$. Let us assume that $G$ is not Berge and consider the following:
	\begin{itemize}
		\item If $P$ is even, then $v_1, v_2, v_3, v_4, P, v_1$ is an odd hole,
		\item If $P$ is of length 3. \todo{I merged a couple of proofs from \cite{MC06}, check in the morning if this is correct.} Let us name its vertices $v_1, p_2, p_3, v_4$. It follows from \cref{lem:Roussel-Rubio}, that there exists an odd antipath between $p_2$ and $p_3$ with interior in $X$. We can complete it with $v_2p_2$ and $v_2p_3$ into an odd antihole.
		\item If $P$ is odd with the length of at least 5 \todo{check in the morning}, it follows from \cref{lem:Roussel-Rubio} that we have $x, y \in X$ with only two edges to $P$ being $xp_2$ and $yp_{n-1}$. This gives us an odd hole: $v_2, x, p_2, \ldots, p_{n-1}, y, v_2$.
	\end{itemize}
\end{proof}

@@

	\noindent Output: Returns whether $G$ contains a configuration of type $T_3$ as an induced subgraph.
\end{alg}
\begin{algtext2}

@@

	Then, for each choice of $v_4$ that is adjacent to $v_1$ and not to $v_2$ and $v_5$ and has a neighbor in $F$ (call it $v_6$) and a nonneibhbor in $X$, we test whether there is a vertex $v_3$, adjacent to $v_2, v_4, v_5$ and not to $v_1$, with a nonneibhbor in $X$. If there is such a vertex $v_3$, find $P$ -- a path from $v_6$ to $v_5$ with interior in $F'$ and return that $v_1, \ldots v_6, P, X$ is a configuration of type $\T_3$. If we exhaust our search and find none, report that graph does not contain it.
\end{algtext2}

>>>>

\paragraph{Finding Pyramids}

\begin{alg}[Test if $G$ contains a Pyramid]
	\label{alg:testPyramid}
	Input: A graph $G$.

	\noindent Output: Returns whether $G$ contains a pyramid as an induced subgraph.
\end{alg}
\begin{algorithmic}[1]
	\mProcedure{Contains-Pyramid}{$G$}
	\mForEach{triangle $b_1, b_2, b_3$} \label{line:pyramidTriangle}
		\mForEach{$s_1, s_2, s_3$, such that for $1 \leq i < j \leq 3$, $\{b_i, s_i\}$ is disjoint
			\lsx  from $\{b_j, s_j\}$ and $b_ib_j$ is the only edge between them} \label{line:pyramidTriple}
			\mIf{there is a vertex $a$, adjacent to all of $s_1, s_2, s_3$, and to at most
				\lsx one of $b_1, b_2, b_3$, such that if $a$ is adjacent to $b_i$, then $b_i = s_i$ \lsx}\label{line:a}
				\ls $M \gets V(G) \setminus \{b_1, b_2, b_3, s_1, s_2, s_3\}$
				\mForEach{$m \in M$} \label{line:pyramidSStart}
					\ls $S_1(m) \gets$ the shortest path between $s_1$ and $m$ such that
					\lsx $s_2, s_3, b_2, b_3$ have no neighbors in its interior, if such a 
					\lsx path exists.
					\ls calculate $S_2(m), S_3(m)$ similarly
					\ls $T_1(m) \gets$ the shortest path between $m$ and $b_1$, such that
					\lsx $s_2, s_3, b_2, b_3$ have no neighbors in its interior, if such a
					\lsx path exists
					\ls calculate $T_2(m), T_3(m)$ similarly
				\mEndFor \label{line:pyramidSEnd}

					\mIf{$s_1 = b_1$} \Comment{calculate all possible $P_1$ paths}
						\ls $P_1(b_1) \gets$ the one-vertex path with vertex $b_1$
						\ls \algorithmicforeach ~$m \in M$ \algorithmicdo~ $P_1(m) \gets$ \textsc{undefined}
					\mElse
						\ls $P_1(b_1) \gets $ \textsc{undefined}
						\mForEach{$m \in M$}
							\mIf{ $m$ is nonadjacent to all of $b_2, b_3, s_2, s_3$ \AND
								\lsx $S_1(m)$ and $T_1(m)$ both exist \AND
								\lsx $V(S_1(m) \cap T_1(m)) = \{m\}$ \AND
								\lsx there are no edges between $V(S_1(m) \setminus m)$ 
								\lsx and $V(T_1(m) \setminus m)$ 
								\lsx
							}
								\ls $P_1(m) \gets s_1-S_1(m)-m-T_1(m)-b_1$
							\mElse
								\ls $P_1(m) \gets$ \textsc{undefined}
							\mEndIf
						\mEndFor
					\mEndIf

				\ls assign $P_2$ and $P_3$ in a similar manner
				\ls $good\_pairs_{1,2} \gets \emptyset$ \Comment{see below for definition}
				\mForEach{$m_1 \in M \cup \{b_1\}$} \Comment{calculate good $(1, 2)$-pairs}
					\mIf{$P_1(m_1) \neq \UND$}
						\ls color black the vertices of $M$ that either belong to \label{line:pyramidColor}
						\lsx $P_1(m_1)$ or have a neighbor in $P_1(m_1)$
						\lsx color all other vertices white.
						\mForEach{$m_2 \in M \cup \{b_2\}$}
							\mIf{$P_2(m_2)$ exists and contains no black vertices} \label{line:pyramidColor2}
								\ls add $(m_1, m_2)$ to $good\_pairs_{1,2}$
							\mEndIf
						\mEndFor
					\mEndIf
				\mEndFor
				\ls calculate $good\_pairs_{1,3}$ and $good\_pairs_{2,3}$ in similar way
				\mForEach{triple $m_1, m_2, m_3$ such that $m_i \in M \cup \{b_i\}$} \label{line:m1m2m3}
					\mIf{$\forall_{1\leq i < j \leq 3}$: $(m_i, m_j)$ is a good $(i, j)$-pair}
						\ls \RETURN \TRUE \label{line:pyramidTrue}
					\mEndIf
				\mEndFor
			\mEndIf \label{line:pyramidEnd}
		\mEndFor
	\mEndFor
	\ls \RETURN \FALSE
	\mEndProcedure
\end{algorithmic}

With definitions as above, for $1 \leq i < j \leq 3$, we say that $(m_i, m_j)$ is a \emph{good $(i, j)$-pair}, iff $m_i \in M \cup \{b_i\}$, $m_j \in M \cup \{b_j\}$, $P_i(m_i)$ and $P_j(m_j)$ both exist, and the sets $V(P_i(m_i))$,$V(P_j(m_j))$ are both disjoint and $b_ib_j$ is the only edge between them.

It is easy to see, that if \textsc{Contains-Pyramid($G$)} outputs that $G$ contains a pyramid, it indeed does -- when we return in line \ref{line:pyramidTrue} $a$ found in line \ref{line:a} can be linked into a triangle $b_1, b_2, b_3$ it was found for via paths $P_1(m_1), P_2(m_2), P_3(m_3)$ for $m_1$, $m_2$, $m_3$ from line \ref{line:m1m2m3}. The proof of the converse is rather technical and we refer to \cite{MC05} for it.

Now we will proove the time complexity.

\begin{theorem}
	Procedure \textsc{Contains-Pyramid($G$)} works in $O(|V|^9)$ time.
\end{theorem}
\begin{proof}
	There are $O(|V|^3)$ triangles (line \ref{line:pyramidTriangle}) and $O(|V|^3)$ triples $s_1$, $s_2$, $s_3$ (line \ref{line:pyramidTriple}), so lines \ref{line:a}-\ref{line:pyramidEnd} are executed at most $O(|V|^6)$ times.

	Checking if there exists an appropriate $a$ takes linear time (line \ref{line:a}). Calculating paths $S_i$ and $T_i$ (lines \ref{line:pyramidSStart}-\ref{line:pyramidSEnd}) takes $O(|V|^2)$ time for each $m \in M$ and $O(|V|^3)$ in total. Similarly, it takes $O(|V|^3)$ time to calculate all $P_i$ paths.

	Then, for each $m_1$ we do at most $O(|V|)$ work in line \ref{line:pyramidColor} and for each $(m_1, m_2)$ we do at most $O(|V|)$ work in line \ref{line:pyramidColor2}.

	Finally, there are $O(|V|^3)$ pairs $m_1, m_2, m_3$ and checking each takes $O(1)$ time.
\end{proof}

@@

\begin{theorem}
	If $G$ contains configuration of type $\T_2$ then $G$ is not Berge.
\end{theorem}
\begin{proof}
	Let $(v_1, v_2, v_3, v_4, P, X)$ be a configuration of type $\T_2$. Let us assume that $G$ is not Berge and consider the following:
	\begin{itemize}
		\item If $P$ is even, then $v_1, v_2, v_3, v_4, P, v_1$ is an odd hole,
		\item If $P$ is of length 3. \todo{I merged a couple of proofs from \cite{MC06}, check in the morning if this is correct.} Let us name its vertices $v_1, p_2, p_3, v_4$. It follows from \cref{lem:Roussel-Rubio}, that there exists an odd antipath between $p_2$ and $p_3$ with interior in $X$. We can complete it with $v_2p_2$ and $v_2p_3$ into an odd antihole.
		\item If $P$ is odd with the length of at least 5 \todo{check in the morning}, it follows from \cref{lem:Roussel-Rubio} that we have $x, y \in X$ with only two edges to $P$ being $xp_2$ and $yp_{n-1}$. This gives us an odd hole: $v_2, x, p_2, \ldots, p_{n-1}, y, v_2$.
	\end{itemize}
\end{proof}

@@

	\noindent Output: Returns whether $G$ contains a configuration of type $T_3$ as an induced subgraph.
\end{alg}
% \begin{algorithmic}
% 	\mProcedure{Contains-T3}{$G$}
% 	\mForEach{$v_1, v_2, v_5 \in V(G)$, so that $v_1v_2$ is an edge and $v_1v_5, v_2v_5$ are nonedges}
% 	\mEndFor
% 	\mEndProcedure
% \end{algorithmic}
% \begin{algorithmic}[1]
% 	\mProcedure{Contains-Pyramid}{$G$}
% 	\ls \RETURN \FALSE
% 	\mEndProcedure
% \end{algorithmic}

% \begin{algtext2}

@@

	Then, for each choice of $v_4$ that is adjacent to $v_1$ and not to $v_2$ and $v_5$ and has a neighbor in $F$ (call it $v_6$) and a nonneibhbor in $X$, we test whether there is a vertex $v_3$, adjacent to $v_2, v_4, v_5$ and not to $v_1$, with a nonneibhbor in $X$. If there is such a vertex $v_3$, find $P$ -- a path from $v_6$ to $v_5$ with interior in $F'$ and return that $v_1, \ldots v_6, P, X$ is a configuration of type $\T_3$. If we exhaust our search and find none, report that graph does not contain it.
% \end{algtext2}
